### 2017-01-11 robust-software.tex
Added emphasis and citations for issue #29 reproducibility; emphasized the prototype nature of scientific software

@@

Many researchers have found out the hard way that there's a world of difference
between ``works for me on my machine'' and ``works for other people on theirs''.
Many common challenges can be avoided by following a few simple rules; doing so
not only accelerates research, but also improves reproducibility.

@@

Being unable to
replicate results is so common that one publication refers to it as "a rite of
passage"~\cite{baker2016}. 
The root cause of this problem is that most research software
isn't \emph{robust}. The difference between running and being
robust is the difference between ``works for me on my machine'' and
``works for other people on a cluster I've never used''. The lack of
robustness also leads to many duplicated efforts,
which slows the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools to perform the same task:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

The good news is,
you don't need to be a professionally-trained programmer to write robust software.
In fact,
some of the best, most reliable pieces of software in many scientific
communities are written by researchers~\cite{prabhu2011,sanders2008}
who have adopted strong software
engineering approaches, have high standards of reproducibility, use good testing
practices, and foster inter-institutional collaborations that grow into strong user
bases. In the bioinformatics community, Bioconductor and Galaxy follow this
path~\cite{gentleman2004,afgan2016}.

@@

\begin{itemize}
\item
  it can be installed on more than one computer with relative ease,
\item
  it works as advertised, and
\item
  it can be integrated with other tools.
\end{itemize}

Our rules are generic and can be applied to all languages, libraries, packages,
documentation styles, and operating systems for both closed-source and open-source software.  
They are also necessary steps toward making computational research reproducible:
after all,
if the tools and libraries used to produce a result cannot be run by others,
they cannot verify or build on published results.

@@

\begin{itemize}

@@

\end{itemize}

>>>>

\usepackage{color}

\newenvironment{reviewed}
{ \color{orange} }

@@

Many researchers have found out the hard way that there's a world of difference
between ``works for me on my machine'' and ``works for other people on theirs''.
Many common challenges can be avoided by following a few simple rules; doing so
not only improves reproducibility, but can accelerate research.

@@

Being unable to
replicate results is so common that one publication refers to it as ``a rite of
passage''~\cite{baker2016}. 
The root cause of this problem is that most research software
is essentially a prototype, and therefore is not 
\emph{robust}. The lack of
robustness in published, distributed software
leads to duplicated efforts with little practical benefit,
which slows the pace of research~\cite{prabhu2011,lawlor2015}. 
Not all software needs to be robust enough to run for other 
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools to perform the same task:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

The good news is,
you don't need to be a professionally-trained programmer to write robust software.
In fact,
some of the best, most reliable pieces of software in many scientific
communities are written by researchers~\cite{prabhu2011,sanders2008}
who have adopted strong software
engineering approaches, have high standards of reproducibility, use good testing
practices, and foster inter-institutional collaborations that grow into strong user
bases. In the bioinformatics community, Bioconductor and Galaxy follow this
path~\cite{gentleman2004,afgan2016}.
Not all scientific software needs to be robust~\cite{varoquaux2015}, but if you
publish a paper about your software, it should, at minimum, satisfy these
rules.

@@

\begin{itemize}
\item
  it can be installed on more than one computer with relative ease,
\item
  it works consistently as advertised, and
\item
  it can be integrated with other tools.
\end{itemize}

Our rules are generic and can be applied to all languages, libraries, packages,
documentation styles, and operating systems for both closed-source and open-source software.  
They are also necessary steps toward making computational research replicable and reproducible:
after all,
if your tools and libraries cannot be run by others,
they cannot be used to verify your results or as a stepping stone for future work~\cite{brown2013}.

Note: practices \emph{not} included in the list below include version control,
code review, unit testing, and other practices essential to maintainability and
sustainable software development.
These elements are essential to building software of any kind
\cite{wilson2014,wilson2016}.
However,
our focus in this short guide is on the software that is already built,
not on the process used to build it.

@@

@@
